% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.


\chapter{Introduction}
\begin{comment}
Chapter plan:
-Context: introduce CDP, explain the need for the VR collaboration component
-Problem description: explain that virtual working environment provides only limited information from the workspace (citing works on Workspace Awareness). Example/Motivation: giant buildings moving at and through you

-Research question??

-Hypothesis: hypothesise that additional auditory cues promote workspace awareness in collaborative VR
-Thesis Overview: provide a high-level overview of the contents of this thesis (Chapter 1 is about.., Chapter 2 is about..)
\end{comment}

% Context
% About CDP Gerhard
\gls{cdp} is a multidisciplinary project at the Technical University of Munich developed by the Chair of Architectural Informatics from the Department of Architecture, the Chair of Augmented Reality from the Department of Informatics, and the Leibniz-Rechnenzentrum M{\"u}nchen.
%About \gls{vr} Sketching Component Sofia
At its core, \gls{cdp} is a multi-touch table \cite[p.~5]{lampe_cdp//vr-sketching_2017} that provides a lot of  functionality useful in urban planning and design (i.e. the ability to interactively add new buildings at a given site on the map, sketch on the \gls{vb}s, run solar envelope and wind simulations, etc.). One of the latest  development directions for the table is a \gls{vr} component, which aims to provide the existing 3D sketching capabilities in an immersive \gls{vr} environment. With the addition of the this component, comes a question of implementing a \gls{cve} to allow multiple people to use it simultaniously.


%Motivation
%Lead to \gls{cdp} \gls{vr} Collaboration component
 In this \gls{cve} users would have to interact with real-sized \gls{vb}s, which points out the initial concern and the motivation for this work. Imagine 2 users performing their separate tasks in the same \gls{cve}: while User 1 is occupied with sketching on the \gls{vb}s and going through the new ideas for an urban district, User 2 is busy with positioning the buildings in the environment to be more esthetically pleasing. The main question was - how would the moving buildings be perceived by a user that is ocuppied with their own task (for example, in case when a \gls{vb} moves through User 1), and how would this influence the quality of collaboration and the satisfaction with the experience in general.  
 

 %Problem description
 This brings up a problem of awareness different users have of each other in the system. \cite{gutwin_descriptive_2002} argues that "... workspace awareness is much harder to maintain in groupware workspaces than in face-to-face environments, and it is often difficult or impossible to determine who else is in the workspace, where they are working, and what they are doing". Authors point out that this is due to certain properties that any virtual workspace, and collaborative software (groupware) in particular have. This is mostly due to the fact that these systems provide only limited information about their current state to the user \cite[p.~414-415]{gutwin_descriptive_2002}. This comes as no surprise, because it would be virtually impossible and impractical to implement all the intricate details of the real world in software that tries to automate it. 
  

 %Hypothesis 
 To overcome these limitations, a user must be provided with the essential information about the environment. In this project, I explore what effect additional auditory cues have on the \gls{wa} of the collaborators. This work extends the \gls{wa} study presented in \cite{gutwin_chalk_2011} by putting ephasis on \gls{wa} in an immersive \gls{vr} environment.
 

% Overview
\paragraph{}
In the Literature Review chapter of this thesis I will explore how we as humans process infomation from the outside world, different channels available for this information. I will also introduce auditory cues as a useful tool for signling and monitoring. Then I turn to \gls{sa}, which can be shortly described as an up-to-date understanding of your current environment. Lastly, we will take a look at subfiled of \gls{sa} - \gls{wa}, and familiarize ourselves with \gls{wa} framework and the \gls{wa} study done by Gutwin et al. in \cite{gutwin_chalk_2011}. The \gls{wa} in Immersive Virtual Reality chapter will present the main user study conducted for this thesis and its results. In the summary ...


% Prev. Overview
\begin{comment}
In Introduction I will write about the main motivation for this work, which is extending the CDP with an immersive collaborative virtual reality component, and the main problems that come along with it, specifically, user awareness of each other's actions. Then I would state my initial hypothesis about the additional cues being able to make the experience more "comfortable".

The Literature Review section of the thesis investigates human information processing and presents auditory cues as a useful tool for signaling and monitoring [1]. Then I turn to the Situation Awareness (SA) in VR, and how it is a measure of a user being able to perceive, interpret, and project the things happening within the environment [2]. We will see that situation awareness has a subfield called Workspace Awareness (WA), which deals exactly with the awareness users have of each other in groupware systems [3,4]. We will also discover that WA deals mainly with monitoring and secondary tasks, and how "comfortable experience" can be described in terms of WA framework. Last part of Literature Review section turns to the field of sonification to explore how it can help with monitoring activities, and the different types and properties of sound that could be unitized [5].

Concept/Discussion. Here I will state the purpose of the project and explain how I am going to solve the problem of WA in the collaborative architectural environment. I will classify the system according to WA framework, and then state that we want to answer the what and where WA questions with auditory cues and (possibly visual cues in the form of a radar view). I will mention the research by Gutwin (2011) and say that I will try to adapt it to our use case. Then I will describe what is needed to do it (what we need to study, experiment with, etc.) and what is needed to get the answers to my research question - whether the additional auditory (and visual cues) help with WA.

Experiments. This section will one by one introduce the experiments made in this research: introduction (with statement of purpose), assumptions (limitations), apparatus and process, results and discussion. Here we will have all pilots in 1 subsection, and all the proper experiments (sound spatial judgement and, possibly, WA study) will have their own subsection.

% Might be the same as summary, except for the future work
In the Conclusions section I will aggregate the all findings and prior discussion to try and give an answer to the research question. I will mention the limitations of the conducted study. This section will end with a look into the future work. 

In Summary I will go over the whole work step by step, recall what has been done, mention the results of the experiments
\end{comment}

%\chapter{Collaborative Virtual Environments}
\chapter{Literature review}
\section{Human Information Processing}
\section{Awareness}
\subsection{Situation Awareness}
\subsection{Workspace Awareness}
\section{Sonification}
\chapter{Concept}
\section{Related Work: Chlak sound}
\paragraph{Intro}
\paragraph{Gutwin vs our study}

\begin{table}[]
  \caption{Study comparison}
  \label{table:study_comp}
  \begin{tabular}{|l|l|l|}
  \hline
                             & Gutwin                & This study           \\ \hline
  Primary (distraction) task & Same as the secondary & Contextually related \\ \hline
  Working (sound) plane      & ?Vertical             & Horizonatal          \\ \hline
  Sound type                 & Auditory icons        & Auditory icons       \\ \hline
  Object of analysis         & Workspace awareness   & Workspace Awareness  \\ \hline
  \end{tabular}
\end{table}

\paragraph{Summary}
\chapter{Experiments}
\section{Pilot studies}
\section{Spatial sound judgment study}
\section{Workspace awareness study}

\input{chapters/ehtics_WA_experiment.tex}

\chapter{Conclusions}
\section{Study limitations}
% the fact that we only sample the level 1 SA/WA (we won't be going into sampling direction of translation guesses from the participants)
\chapter{Summary}

%TODO
Acknowledgements 
Bibliography






%---------------Reference commands and structires

% \chapter{Introduction}\label{chapter:introduction}
% \section{Motivation: Teamwork is important. Creative thinking implies generation of ideas, collaboration, and communication. Virtual Reality (VR) is a great enhancement for creative thinking tool-set in architecture.}
%\subsection{Solution validation/evaluation in HCI: methods, and principles.}
\begin{comment}
Methodology: approach to solving the problem; chosen HCI methodology for the final evaluation - no idea
a. Chosen HCI evaluation methodology
\end{comment}


\begin{comment}
See~\autoref{tab:sample}, \autoref{fig:sample-drawing}, \autoref{fig:sample-plot}, \autoref{fig:sample-listing}.
\begin{table}[htpb]
  \caption[Example table]{An example for a simple table.}\label{tab:sample}
  \centering
  \begin{tabular}{l l l l}
    \toprule
      A & B & C & D \\
    \midrule
      1 & 2 & 1 & 2 \\
      2 & 3 & 2 & 3 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htpb]
  \centering
  % This should probably go into a file in figures/
  \begin{tikzpicture}[node distance=3cm]
    \node (R0) {$R_1$};
    \node (R1) [right of=R0] {$R_2$};
    \node (R2) [below of=R1] {$R_4$};
    \node (R3) [below of=R0] {$R_3$};
    \node (R4) [right of=R1] {$R_5$};

    \path[every node]
      (R0) edge (R1)
      (R0) edge (R3)
      (R3) edge (R2)
      (R2) edge (R1)
      (R1) edge (R4);
  \end{tikzpicture}
  \caption[Example drawing]{An example for a simple drawing.}\label{fig:sample-drawing}
\end{figure}

\begin{figure}[htpb]
  \centering

  \pgfplotstableset{col sep=&, row sep=\\}
  % This should probably go into a file in data/
  \pgfplotstableread{
    a & b    \\
    1 & 1000 \\
    2 & 1500 \\
    3 & 1600 \\
  }\exampleA
  \pgfplotstableread{
    a & b    \\
    1 & 1200 \\
    2 & 800 \\
    3 & 1400 \\
  }\exampleB
  % This should probably go into a file in figures/
  \begin{tikzpicture}
    \begin{axis}[
        ymin=0,
        legend style={legend pos=south east},
        grid,
        thick,
        ylabel=Y,
        xlabel=X
      ]
      \addplot table[x=a, y=b]{\exampleA};
      \addlegendentry{Example A};
      \addplot table[x=a, y=b]{\exampleB};
      \addlegendentry{Example B};
    \end{axis}
  \end{tikzpicture}
  \caption[Example plot]{An example for a simple plot.}\label{fig:sample-plot}
\end{figure}

\begin{figure}[htpb]
  \centering
  \begin{tabular}{c}
  \begin{lstlisting}[language=SQL]
    SELECT * FROM tbl WHERE tbl.str = "str"
  \end{lstlisting}
  \end{tabular}
  \caption[Example listing]{An example for a source code listing.}\label{fig:sample-listing}
\end{figure}
\end{comment}